\section{Complex Numbers}
The complex numbers are another set of numbers, which encompass all the numbers see so far:

\begin{itemize}
	\item $\mathbb{N}$ represents the natural numbers, or positive integers.
	\item $\mathbb{Z}$ represents the integers.
	\item $\mathbb{Q}$ represents the rational numbers, or all numbers which can be expressed as fractions.
	\item $\mathbb{R}$ represents the real numbers, which includes irrational numbers such as $\pi$ and $e$.
\end{itemize}

All of these sets include all of the ones above them. In a similar way, the complex numbers, $\mathbb{C}$, include all of these sets. Complex numbers can be thought of on a 2-dimensional number line with a real axis and an imaginary axis.

\subsection{The complex unit}
The complex unit, $i$, is equal to $\sqrt{-1}$. In this fashion:

\begin{ea}[rCl]
	\sqrt{-25} & = & \sqrt{-1 \times 25}
	\nonumber\\
			   & = & \sqrt{-1} \times \sqrt{25}
	\nonumber\\
			   & = & 5i
	\nonumber\\
	\sqrt{-100} & = & 10i
	\nonumber\\
	i^2 & = & -1
\end{ea}

\subsection{Writing complex numbers}
Complex numbers, usually represented by $z$, can be expressed in the form $a+bi$, where $a,b \in \mathbb{R}$. In this form then $a$ is referred to as the real part, while $b$ is called the imaginary part. They can have various operations done on them in a logical way:
\begin{ea}[rCl]
	(3+4i)+(5-3i) & = & 8+i
	\nonumber\\
	(4+2i)(3-i) & = & 12 + 2i - 2i^2
	\nonumber\\
				& = & 14 + 2i
\end{ea}

\subsection{Realising the denominator}
It's possible to multiply two complex numbers together and have them give an answer which only has a real part. This can be useful when there's a complex number on the bottom of a fraction, or simply used to express the division of complex numbers nicely:

\begin{ea}[rCl]
	1 \div (1-2i)
	\nonumber\\
	& = & \frac{1}{1-2i}
	\nonumber\\
	& = & \frac{1(1+2i)}{(1-2i)(1+2i)}
	\nonumber\\
	& = & \frac{1+2i}{5}
	\nonumber\\
	& = & \frac{1}{5} + \frac{2}{5}i
\end{ea}

\subsection{Complex conjugates}
The complex conjugate of a number is the same number, with the imaginary part multiplied by -1. So if $z=a+bi$, then its complex conjugate $z^*=a-bi$. These together are referred to as a conjugate pair. They can be useful because $z+z^*=2a$, and $zz^*=a^2+b^2$, both of which use an operation to give a real result out of the two imaginary numbers.

\subsection{Complex powers}
The powers of $i$ go in a cycle:
\begin{ea}[rCl]
	i^1 & = & i
	\nonumber\\
	i^2 & = & -1
	\nonumber\\
	i^3 & = & -i
	\nonumber\\
	i^4 & = & 1
	\nonumber\\
	i^5 & = & i
\end{ea}
This cycle has a period of 4, and so we can divide any power of $i$ by 4 and use the remainder to see what the power of $i$ it's equal to.

\subsection{Imaginary roots of quadratics}
For any quadratic $f(z)$, then if $z_1$ is a root, then $z_1^*$ must necessarily also be a root. One convention is to let $z_1=\alpha$, and $z_1^*=\beta$. Then:

\begin{ea}[rCl]
	(z-\alpha)(z-\beta) & = & 0
	\nonumber\\
	z^2 -\alpha z - \beta z + \alpha \beta & = & 0
	\nonumber\\
	z^2 - z(\alpha + \beta) + \alpha \beta & = & 0
\end{ea}

Which means to represent the quadratic in the form $ax^2+bx+c$ then $b=\alpha + \beta$ and $c=\alpha \beta$.

\subsection{Cubics and quartics with complex roots}
For a quadratic, there are three possible options for the roots being real or not. There could be 2 real roots. There could be 1 repeated root. Or there could be 2 complex roots, one conjugate pair. For cubics, there must always be at least one real root, because there can't ever be an odd number of complex roots. If you have the real root, you can algebraically divide to give a quadratic with the two complex roots.

This oncept extends to quartics where there could be 4 real roots, 2 real and 2 complex, 4 complex roots, and so on with repititions.
